---
title: "The Comprehenssive R Learning Pathway"
author: "Dr Nikhil Chandra Sarkar, Founder of Data Simulation Research Lab"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Let's check the current working directory on your computer

```{r}
getwd()

```
# Remove all the memory from workspace in R Environment

```{r}
rm(list = ls())

```


# Install a package

```{r}
#install.packages('tidyverse')

```

# Load installed library

```{r}
library(tidyverse)  # For data cleaning and tidy

```


# Load other installed libraries

```{r}
library(data.table)     # For data frame
library(lubridate)      # For date and time data
library(ggplot2)        # For visualisation

```

## Let's get strated with a datset from a package (ggplot2)

ggplot2::mpg is a Fuel economy data from 1999 to 2008 for 38 popular models of cars.

```{r}
data <- mpg
data %>% head(10)

```

## Get quick data summary

```{r}
data %>%
  summary()

```

## Most basic opertions often used in data transformation: 

To learn the five key dplyr functions that allow you to solve the vast majority of your data manipulation challenges:

1. Pick observations by their values: filter()
2. Reorder the rows: arrange()
3. Pick variables by their names: select()
4. Create new variables with functions of existing variables: mutate()
5. Collapse many values down to a single summary: summarise()

## Use filter function

```{r}
library(dplyr)
# Filter the mpg data for model "a4" or "malibu"

filter(data, model =="a4" )

filter(data, model =='a4' | model  =='malibu')

# Alternative for above operation
data %>% filter(model=="a4" | model =="malibu")

```

## Use summarise function

```{r}
# descripitve statics: 
data %>% filter( model=="a4" | model =="malibu" ) %>% 
  group_by(model) %>% 
  summarise (Count=n())

```

## Use arrange function

```{r}
# Rearrange mpg data based on year in decendening order of year
arrange(data, +year) 

# Rearrange mpg data based on year in accending order of year
arrange(data, -year)

```

## Use select function

```{r}
# Select different variables from the dataset
select(data, year, model, displ, trans)

# Alternative way using

data %>% select(year, model, displ, trans)

```


## Generate new variable using existing variables by mutate()

```{r}
# Generate new variables for the mpg data set
mutate(data,  differance  =  cty - hwy)

# Aletrnative way
data %>% mutate(differance  =  cty-hwy)

```
## Use summarise

```{r}
# Generate summary statistics for a varible
summarise(data, count =n(), mean(cty),sd(cty),min(cty), max(cty))

# Aletrnative way

data %>% summarise(count =n(), mean(cty),sd(cty),min(cty), max(cty))

# Descriptive statistics
data %>% group_by(displ) %>% summarise(count =n())
data %>% group_by(displ) %>% summarise(count =n(), mean(displ), sd(displ), min(displ), max(displ))

```

## Visualisation

```{r}
library(ggplot2)
ggplot() +
  layer(
    data = diamonds, mapping = aes(x = carat, y = price),
geom = "point", stat = "identity", position = "identity" ) +
  scale_y_continuous() +
  scale_x_continuous() +
  coord_cartesian()

```
## Visualisation

```{r}

  library(ggplot2)
  ggplot() +
  layer(
  data = mpg, mapping = aes(x = displ, y = cty),
  geom = "point", stat = "identity", position = "identity" ) +
  scale_y_continuous() +
  scale_x_continuous() +
  coord_cartesian() +
  xlab("Engine displacement (litres)") +
  ylab("City miles per gallon")
```
